Reproducibility of COVID-19 pre-prints

### 2022-01-13 outputs/paper/paper.Rmd
Address some more comments

---
title: "Reproducibility of COVID-19 pre-prints"
author: 
  - Annie Collins^[University of Toronto.]
  - Rohan Alexander^[University of Toronto, rohan.alexander@utoronto.ca.]
thanks: "**Preprint:** This present manuscript is a version of a preprint by the authors [@collins2021reproducibility]. The text, tables and figures of this manuscript overlap with those of the preprint version by the authors. **Acknowledgements:** We thank Amy Farrow, Jessica Gronsbell, Monica Alexander, and Thomas William Rosenthal for helpful comments. Two anonymous reviewers and the editor provided extensive comments that substantially improved this paper and we are grateful to them. **Funding statement:** Collins thanks CANSSI Ontario and Monica Alexander for financial support. **Data accessibility statement:** Code and data are available at: https://github.com/anniecollins/reproducibility_markers_in_covid19_preprints. **Author contributions:** Collins had the original idea and wrote the first draft of the paper. Collins and Alexander obtained and created the datasets. Collins and Alexander analysed the data. All authors interpreted the data and contributed writing the paper and approved the final version."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "To examine the reproducibility of COVID-19 research, we create a dataset of pre-prints posted to arXiv, bioRxiv, medRxiv, and SocArXiv between 28 January 2020 and 30 June 2021 that are related to COVID-19. We extract the text from these pre-prints and parse them looking for keyword markers signalling the availability of the data and code underpinning the pre-print. For the pre-prints that are in our sample, we are unable to find markers of either open data or open code for 75 per cent of those on arXiv, 67 per cent of those on bioRxiv, 79 per cent of those on medRxiv, and 85 per cent of those on SocArXiv. We conclude that there may be value in having authors categorize the degree of openness of their pre-print as part of the pre-print submissions process, and more broadly, there is a need to better integrate open science training into a wide range of fields."
output:
  bookdown::pdf_document2:
    keep_tex:  true
toc: FALSE
bibliography: references.bib
---

@@

```{r loadpackages, include=FALSE}
library(here)
library(knitr)
library(lubridate)
library(tidyverse)
library(stargazer)
# library(wesanderson)
library(kableExtra)
library(DiagrammeR)
```

@@

socarxiv_open_data_results <- 
  read_csv(here::here("outputs/data/socarxiv_open_data_results.csv"))
socarxiv_2019_open_data_results <- 
  read_csv(here::here("outputs/control-data/socarxiv_2019_open_data_results.csv"))
socarxiv_metadata_2019 <- 
  read_csv(here::here("outputs/control-data/socarxiv_metadata_2019.csv"))

@@

Scientists use open repositories of papers to disseminate their research more quickly than is possible in traditional journals or conference proceedings, and also to obtain feedback on their work prior to publication. These repositories, such as arXiv, bioRxiv, medRxiv, and SocArXiv, are a critical component of scientific communication and a lot of research builds on the pre-prints posted there. Pre-print repositories have been especially important during the 2019 novel coronavirus (COVID-19) pandemic and the changes it has imposed on the scientific community [@else2020]. The centrality of pre-prints to science means that it is important that the results that are posted are credible. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a pre-print.

Neither peer-review nor credentials are a panacea nor a guarantee of quality. And the gate-keeping and slow publication times of traditional journals mean pre-print repositories are important. But it is important that scientists impose standards on themselves, and arguably repositories have a role to play here. Following @weissgerber2021automated, we examine pre-prints about COVID-19 posted to arXiv, bioRxiv, medRxiv, and SocArXiv from 28 January 2020 through to 30 June 2021. By way of background, each of these four repositories has a different focus: arXiv is general although it has especially high rates of usage from fields like mathematics, physics, and computer science, bioRxiv focuses on biological sciences, medRxiv focuses on health sciences, and SocArXiv focuses on social sciences.

We search for markers of open science as indicators of reproducibility, specifically open data and open code. The definition of reproducibility tends to vary by context and academic field. For the purposes of this paper, we define reproducibility to mean the ability for different researchers to achieve the same results given the same data and computational methods as the original source. This contrasts replicability, which we define as the ability for different researchers to achieve consistent results by conducting the full data collection and analysis process in lieu of reusing original data. These definitions match that of @NAP25303 and @cacioppo2015social.

We find that of the papers sampled, approximately 75 per cent of papers from arXiv, 67 per cent of papers from bioRxiv, 79 per cent of papers from medRxiv, and 85 per cent of papers from SocArXiv contain neither open data nor open code markers. A summary of our main results is contained in Figure \@ref(fig:flowchart). Examining trends over time, we find that the proportion of pre-prints containing open data or code markers has fluctuated but shown no obvious trend throughout the pandemic. We also find that the presence of open data or open code markers seems to have little association with a pre-print's subsequent publication, and the subset of sampled pre-prints that have been published contains approximately the same proportion of papers with these markers.

@@

all [label =  'All pre-prints posted between 28 January 2020 and 30 June 2021']
onlycovid [label =  'Those about COVID-19']
arxivcovidcount [label =  'arXiv: 4,250 pre-prints']
biorxivcovidcount [label =  'bioRxiv: 4,146 pre-prints']
medrxivcovidcount [label =  'medRxiv: 13,194 pre-prints']
socarxivcovidcount [label =  'SocArXiv: 478 pre-prints']
randomsample [label =  'Random sampling stratified by repository']
arxivsamplecount [label =  'arXiv: 1,000 pre-prints']
biorxivsamplecount [label =  'bioRxiv: 1,000 pre-prints']
medrxivsamplecount [label =  'medRxiv: 1,500 pre-prints']
socarxivsamplecount [label =  'SocArXiv: 478 pre-prints']
opencodeanddata [label =  'Those with markers of open code or open data']
arxivopencount [label =  'arXiv: 248 pre-prints']
biorxivopencount [label =  'bioRxiv: 326 pre-prints']
medrxivopencount [label =  'medRxiv: 320 pre-prints']
socarxivopencount [label =  'SocArXiv: 73 pre-prints']

# edge definitions with the node IDs
all -> onlycovid -> {arxivcovidcount biorxivcovidcount medrxivcovidcount socarxivcovidcount} -> randomsample -> {arxivsamplecount biorxivsamplecount medrxivsamplecount socarxivsamplecount} -> opencodeanddata -> {arxivopencount biorxivopencount medrxivopencount socarxivopencount}
}")
```

The remainder of this paper is structured as follows: in Section \@ref(methodology) we discuss the process of constructing our dataset through retrieving pre-prints from the arXiv, bioRxiv, medRxiv, and SocArXiv repositories and mining them for open data and open code markers. In Section \@ref(results), we present the results and key findings of this process. Finally, in Section \@ref(discussion) we discuss the implications of these findings in the broader context of reproducibility and science during the COVID-19 pandemic, as well as next steps to expand on our findings and questions raised in the research process.

@@

Our primary dataset consists of pre-print metadata extracted from the arXiv, bioRxiv, medRxiv, and SocArXiv pre-print repositories via their respective Application Programming Interfaces (APIs). This metadata varies by repository, but generally includes the following information: title, abstract, author(s), date created, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, published DOI (if the pre-print has since been published in a peer-reviewed journal), and download link. The data collection process was conducted separately for COVID-19 and pre-COVID-19 papers.

For COVID-19-related pre-prints, we first created a local copy of each repository containing all metadata for pre-prints posted between 1 January, 2020, and 30 June, 2021. In the case of SocArXiv there are fields that are self-reported by the submitting author about the availability of code and data. We classified individual pre-prints as "COVID-19-related" based on whether they contained one or more of the following terms in their title or abstract (case insensitive): "COVID-19", "COVID 19", "corona virus", "coronavirus", "coronavirus-2", "SARS-CoV-2", "SARSCoV-2", and "2019-nCoV". We then randomly sampled pre-prints for further analysis, except for SocArXiv, for which we analyzed all pre-prints that matched the above criteria, because there were only 478.

@@

We checked our sampled pre-prints for open data and code markers using the Open Data Detection in Publications (ODDPub) text mining algorithm [@citePDDPub] within the `oddpub` R package [@citeODDPubpackage]. This required downloading each pre-print as a PDF, which we did using `heapsofpapers` [@citealexandermafouz] and then converting the PDFs to text files. Some papers from SocArXiv were only available in Word document format (.docx) and were downloaded as such and then converted to text format. We then conducted the open data and open code detection procedure, which involved searching for keywords and other markers of open data and open code availability. This was conducted using the `open_data_search()` function from the `oddpub` package. Under original validation process, the ODDPub algorithm performed with a sensitivity of 0.73 and a specificity of 1.00 for open code detection and a sensitivity of 0.73 and a specificity of 0.97 for open data detection compared with manual screening [@citePDDPub].

What constitutes open code or open data is complicated and discipline specific. The details of the `oddpub` approach are available in @citePDDPub. The general criteria are that specific mention should be made of where the data and code are located, and that data should be as close to raw as possible. Data and code must also be freely accessible to anyone (no request, application, registration process, or affiliation required). Our work was conducted using the statistical programming language R [@citeR].

@@

Although SocArXiv's metadata contained author-reported information concerning data and code availability for each pre-print, we elected to use the results of the ODDPub algorithm in our analysis to ensure comparability of results between pre-prints from all repositories. 
<!-- Notable discrepancies between the ODDPub algorithm and authors' self-reported data and code availability are documented in Appendix \@ref(socarxivissues). -->

```{r monthly-papers-total, fig.cap = "Number of pre-prints related to COVID-19 posted to arXiv (N = 4,250), bioRxiv (N = 4,146), medRxiv (N = 13,194), and SocArXiv (N = 478) per month", out.width="90%", fig.align = "center", fig.pos="H"}
just_arxiv <- 
  all_arxiv_results  %>% 
  select(submitted) %>% 
  mutate(type = "arXiv") %>% 
  rename(date = submitted)

@@

just_socarxiv <- 
  socarxiv_open_data_results  %>% 
  select(date_created) %>%
  rename(date = "date_created") %>%
  mutate(date = as.Date(date), type = "SocArXiv")

all <- rbind(just_arxiv, just_medRxiv, just_bioRxiv, just_socarxiv)

@@

```{r 2019-results-med-arxiv, out.width="90%", fig.align='center', fig.cap = "Number of pre-prints posted per month in 2019 to arXiv (N = 150,018), bioRxiv (N = 31,752), medRxiv (N = 913), and SocArXiv (N = 1,614) per month", fig.pos="H"}
bio_2019 <- bio_2019_results %>%
  select(date) %>%
  mutate(type = "bioRxiv")

@@

socarxiv_2019 <- socarxiv_metadata_2019 %>%
  select(date_created) %>%
  mutate(type = "socArXiv") %>%
  rename(date = date_created)

@@

all_2019 <- rbind(arxiv_2019, bio_2019, med_2019, socarxiv_2019)

@@

Between June and December 2019, the number of pre-prints posted to medRxiv monthly saw an overall increase, which may be expected as the repository gained recognition and popularity in the medical research community (Figure \@ref(fig:2019-results-med-arxiv)). The number of pre-prints posted monthly to bioRxiv also saw a slight overall increase throughout 2019, while the number of those posted to arXiv and SocArXiv fluctuated throughout the year (Figure \@ref(fig:2019-results-med-arxiv)). Due to its relative immaturity at the beginning of the COVID-19 pandemic, a significant portion of medRxiv's overall usage has been dedicated to COVID-19-related research. In total, 21,647 pre-prints were posted to medRxiv between June 2019 and 30 June, 2021, 13,194 of which (approximately 61 per cent) relate to COVID-19.

Of the analyzed pre-prints from 2019, 93 per cent of those posted to arXiv, 63 per cent of those posted to bioRxiv, 75 per cent of those posted to medRxiv, 92 per cent of those posted to SocArXiv showed no indication of open data or open code (Table \@ref(tab:2019-published)).

```{r monthly-papers-condition-stack-2019, fig.cap = "Number of sampled pre-prints posted in 2019 from arXiv (n = 1,200), bioRxiv (n = 1,200), medRxiv (n = 913), and SocArXiv (n = 1,200) distinguished by presence of open data or code markers", out.width="90%", fig.align = "center"}
# Number of papers for each open data/code status per month
open_status_med_2019 <-
  med_2019_open_data_results %>%
  select(title, date, is_open_code, is_open_data) %>%
  mutate(condition = is_open_data + is_open_code, repo = "medRxiv")

@@

open_status_socarxiv_2019 <-
  socarxiv_2019_open_data_results %>%
  select(title, date_created, is_open_code, is_open_data) %>%
  mutate(condition = is_open_data + is_open_code, repo = "SocArXiv", date = as.character(date_created)) %>%
  select(title, date, is_open_code, is_open_data, condition, repo)

open_status <- rbind(open_status_med_2019, open_status_bio_2019, open_status_arxiv_2019, open_status_socarxiv_2019)

@@

open_status %>%
  ggplot() +
  geom_bar(aes(x = factor(floor_date(date, "month")),
              fill=factor(condition, levels = c("Neither", "Open Code", "Open Data", "Both")),
              position = "stack")) +
  labs(fill = "Type of marker") +
  scale_x_discrete(breaks=c("2019-01-01","2019-07-01", "2019-12-01"),
                   labels=c("Jan 2019", "Jul 2019", "Dec 2019")) +
  ylab("Number of pre-prints") +
  xlab("Month") +
  labs(title="Sampled pre-prints by month, 2019", subtitle="Note the different y-axes.") +
  theme_minimal() +
  scale_fill_manual(values = cbPalette) +
  facet_wrap(~repo,
             scales = "free_y")
```

@@

socarxiv_published_2019 <- 
  socarxiv_metadata_2019 %>% 
  select(title, date_created, links.doi) %>% 
  mutate(repo = "socArXiv", published = as.numeric(!is.na(links.doi))) %>%
  rename(date = date_created) %>%
  select(title, date, published, repo)

@@

published_2019 <- rbind(med_published_2019, bio_published_2019, socarxiv_published_2019, arxiv_published_2019)

@@

socarxiv_2019_published_open_status <- 
  socarxiv_2019_open_data_results %>%
  mutate(published = !is.na(links.doi) %>% as.numeric()) %>%
  select(title, date_created, published, is_open_code, is_open_data) %>%
  rename(date = date_created) %>%
  mutate(condition = is_open_data + is_open_code, repo = "socArXiv")

@@

published_2019_open_status <- rbind(med_2019_published_open_status, bio_2019_published_open_status, socarxiv_2019_published_open_status, arxiv_2019_published_open_status)

@@

published_2019_open_status %>%
  count(repo, published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>%
  mutate(Status = ifelse(published == 0, "Unpublished", "Published"), `Proportion with neither`= round((`Neither`)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  select(Status, Both, Neither, `Open code`, `Open data`, `Proportion with neither`) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published, 2019",
        booktabs = TRUE,
        format.args = list(big.mark = ",")) %>%
  group_rows(index = c("arXiv" = 2, "bioRxiv" = 2, "medRxiv" = 2, "socArXiv" = 2))
```

Examining publication rates for pre-pandemic papers, we observe that 41 per cent of pre-prints posted to arXiv, 64 per cent of pre-prints posted to bioRxiv, 61 per cent of pre-prints posted to medRxiv, and 37 per cent of pre-prints posted to SocArXiv during 2019 were eventually peer reviewed and published (Table \@ref(tab:2019-published)). When disaggregated by open data and code status, we find that published and unpublished pre-prints contain open data and code markers in similar proportions (Table \@ref(tab:2019-open-publishing)).

@@

From the collection of all pre-prints related to COVID-19, we randomly sampled 3,910 pre-prints to analyze, stratified by repository. This sample is broken down as follows: 1,500 from medRxiv, 1,000 from arXiv, 1,000 from bioRxiv, and 478 from SocArXiv. Broadly, we are unable to find markers of either open data or open code for 3,011 pre-prints or approximately 76 per cent of our sample (Table \@ref(tab:summarycounts-repositories)). Of the remaining pre-prints, 7 per cent contained open code markers only, 10 per cent contained open data markers only, and 8 per cent included markers of both open data and open code.

When differentiated by repository, we observe that open data and code markers were absent from 75 per cent of the sampled arXiv pre-prints, 67 per cent of the sampled bioRxiv pre-prints, 79 per cent of the sampled medRxiv pre-prints, and 85 per cent of all COVID-19-related SocArXiv pre-prints. The distribution of the remaining portion of pre-prints also varies by repository (Table \@ref(tab:summarycounts-repositories)). Notably, 28 per cent of sampled pre-prints from bioRxiv contained open data markers and 22 per cent of sampled arXiv pre-prints contained markers of open code, the highest proportions of any repository for each type of marker. Our results are similar to @mcguinness2021descriptive, who focus on medRxiv and find that 23 per cent describe open data.

@@

just_socarxiv <-
  socarxiv_open_data_results  %>%
  select(date_created, is_open_data, is_open_code) %>%
  rename(date = "date_created") %>%
  mutate(type = "SocArXiv")

all <- rbind(just_arxiv, just_medRxiv, just_bioRxiv, just_socarxiv)

@@

The distribution of total sampled pre-prints and sampled pre-prints with open data or code markers roughly follows that of COVID-19-related pre-prints posted in general (Figure \@ref(fig:monthly-papers-condition-stack)). The proportion of pre-prints with open data or code has fluctuated over time but shows no consistent overall increase or decrease throughout the course of the pandemic, nor in conjunction with increases or decreases in the total number of pre-prints posted to any given repository. In our datasets, very few (if any) pre-prints were sampled for the month of January 2020. None of these pre-prints contained open data or open code markers, thus the 0 per cent rate of open data and code for this month across all repositories should be considered an outlier.

<!-- It is also important to note that pre-prints posted during the early months of the pandemic were likely using and reusing publicly available data sources due toan inability to collect original data within a short timeframe. -->

@@

# Tables

@@

all_2019 <- rbind(just_2019_medRxiv, just_2019_bioRxiv, just_2019_socarxiv, just_2019_arxiv)

>>>>

---
title: "Reproducibility of COVID-19 pre-prints"
author: 
  - Annie Collins^[University of Toronto.]
  - Rohan Alexander^[University of Toronto, rohan.alexander@utoronto.ca.]
thanks: "**Preprint:** This present manuscript is a version of a preprint by the authors [@collins2021reproducibility]. The text, tables and figures of this manuscript overlap with those of the preprint version by the authors. **Acknowledgements:** We thank Amy Farrow, Jessica Gronsbell, Monica Alexander, and Thomas William Rosenthal for helpful comments. Two anonymous reviewers and the editor provided extensive comments that substantially improved this paper and we are grateful to them. **Funding statement:** Collins thanks CANSSI Ontario and Monica Alexander for financial support. **Data accessibility statement:** Code and data are available at: https://github.com/anniecollins/reproducibility_markers_in_covid19_preprints. **Author contributions:** Collins had the original idea and wrote the first draft of the paper. Collins and Alexander obtained and created the datasets. Collins and Alexander analysed the data. All authors interpreted the data and contributed writing the paper and approved the final version."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "To examine the reproducibility of COVID-19 research, we create a dataset of pre-prints posted to arXiv, bioRxiv, and medRxiv between 28 January 2020 and 30 June 2021 that are related to COVID-19. We extract the text from these pre-prints and parse them looking for keyword markers signalling the availability of the data and code underpinning the pre-print. For the pre-prints that are in our sample, we are unable to find markers of either open data or open code for 75 per cent of those on arXiv, 67 per cent of those on bioRxiv, and 79 per cent of those on medRxiv. We conclude that there may be value in having authors categorize the degree of openness of their pre-print as part of the pre-print submissions process, and more broadly, there is a need to better integrate open science training into a wide range of fields."
output:
  bookdown::pdf_document2:
    keep_tex:  true
toc: FALSE
bibliography: references.bib
---

@@

```{r loadpackages, include=FALSE}
library(here)
library(knitr)
library(lubridate)
library(tidyverse)
library(stargazer)
# library(wesanderson)
library(kableExtra)
library(DiagrammeR)
library(patchwork)
```

@@

# socarxiv_open_data_results <- 
#   read_csv(here::here("outputs/data/socarxiv_open_data_results.csv"))
# socarxiv_2019_open_data_results <- 
#   read_csv(here::here("outputs/control-data/socarxiv_2019_open_data_results.csv"))
# socarxiv_metadata_2019 <- 
#   read_csv(here::here("outputs/control-data/socarxiv_metadata_2019.csv"))

@@

Scientists use open repositories of papers to disseminate their research more quickly than is possible in traditional journals or conference proceedings, and also to obtain feedback on their work prior to publication. These repositories, such as arXiv, bioRxiv, and medRxiv, are a critical component of scientific communication and a lot of research builds on the pre-prints posted there. Pre-print repositories have been especially important during the 2019 novel coronavirus (COVID-19) pandemic and the changes it has imposed on the scientific community [@else2020]. The centrality of pre-prints to science means that it is important that the results that are posted are credible. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a pre-print.

Neither peer-review nor credentials are a panacea nor a guarantee of quality. And the gate-keeping and slow publication times of traditional journals mean pre-print repositories are important. But it is important that scientists impose standards on themselves, and arguably repositories have a role to play here. Following @weissgerber2021automated, we examine pre-prints about COVID-19 posted to arXiv, bioRxiv, and medRxiv from 28 January 2020 through to 30 June 2021. By way of background, each of these three repositories has a different focus: arXiv is general although it has especially high rates of usage from fields like mathematics, physics, and computer science, bioRxiv focuses on biological sciences, and medRxiv focuses on health sciences.

We search for markers of open science as indicators of reproducibility, specifically open data and open code. The definition of reproducibility tends to vary by context and academic field [@barba2018terminologies]. For the purposes of this paper, we define reproducibility to mean the ability for different researchers to achieve the same results given the same data and computational methods as the original source. This contrasts replicability, which we define as the ability for different researchers to achieve consistent results by conducting the full data collection and analysis process in lieu of reusing original data. These definitions match that of @NAP25303 and @cacioppo2015social.

We find that of the papers sampled, approximately 75 per cent of papers from arXiv, 67 per cent of papers from bioRxiv, and 79 per cent of papers from medRxiv contain neither open data nor open code markers. A summary of our main results is contained in Figure \@ref(fig:flowchart). Examining trends over time, we find that the proportion of pre-prints containing open data or code markers has fluctuated but shown no obvious trend throughout the pandemic. We also find that the presence of open data or open code markers seems to have little association with a pre-print's subsequent publication, and the subset of sampled pre-prints that have been published contains approximately the same proportion of papers with these markers.

@@

all [label =  'All pre-prints posted between 28 January 2020 and 30 June 2021']
onlycovid [label =  'Those about COVID-19']
arxivcovidcount [label =  'arXiv: 4,250 pre-prints']
biorxivcovidcount [label =  'bioRxiv: 4,146 pre-prints']
medrxivcovidcount [label =  'medRxiv: 13,194 pre-prints']
# socarxivcovidcount [label =  'SocArXiv: 478 pre-prints']
randomsample [label =  'Random sampling stratified by repository']
arxivsamplecount [label =  'arXiv: 1,000 pre-prints']
biorxivsamplecount [label =  'bioRxiv: 1,000 pre-prints']
medrxivsamplecount [label =  'medRxiv: 1,500 pre-prints']
# socarxivsamplecount [label =  'SocArXiv: 478 pre-prints']
opencodeanddata [label =  'Those with markers of open code or open data']
arxivopencount [label =  'arXiv: 248 pre-prints']
biorxivopencount [label =  'bioRxiv: 326 pre-prints']
medrxivopencount [label =  'medRxiv: 320 pre-prints']
# socarxivopencount [label =  'SocArXiv: 73 pre-prints']

# edge definitions with the node IDs
all -> onlycovid -> {arxivcovidcount biorxivcovidcount medrxivcovidcount} -> randomsample -> {arxivsamplecount biorxivsamplecount medrxivsamplecount} -> opencodeanddata -> {arxivopencount biorxivopencount medrxivopencount}
}")
```

The remainder of this paper is structured as follows: in Section \@ref(methodology) we discuss the process of constructing our dataset through retrieving pre-prints from the arXiv, bioRxiv, and medRxiv repositories and mining them for open data and open code markers. In Section \@ref(results), we present the results and key findings of this process. Finally, in Section \@ref(discussion) we discuss the implications of these findings in the broader context of reproducibility and science during the COVID-19 pandemic, as well as next steps to expand on our findings and questions raised in the research process.

@@

Our primary dataset consists of pre-print metadata extracted from the arXiv, bioRxiv, and medRxiv pre-print repositories via their respective Application Programming Interfaces (APIs). This metadata varies by repository, but generally includes the following information: title, abstract, author(s), date created, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, published DOI (if the pre-print has since been published in a peer-reviewed journal), and download link. The data collection process was conducted separately for COVID-19 and pre-COVID-19 papers.

For COVID-19-related pre-prints, we first created a local copy of each repository containing all metadata for pre-prints posted between 1 January, 2020, and 30 June, 2021. 
<!-- In the case of SocArXiv there are fields that are self-reported by the submitting author about the availability of code and data.  -->
We classified individual pre-prints as "COVID-19-related" based on whether they contained one or more of the following terms in their title or abstract (case insensitive): "COVID-19", "COVID 19", "corona virus", "coronavirus", "coronavirus-2", "SARS-CoV-2", "SARSCoV-2", and "2019-nCoV". We then randomly sampled pre-prints for further analysis.
<!-- , except for SocArXiv, for which we analyzed all pre-prints that matched the above criteria, because there were only 478. -->

@@

We checked our sampled pre-prints for open data and code markers using the Open Data Detection in Publications (ODDPub) text mining algorithm [@citePDDPub] within the `oddpub` R package [@citeODDPubpackage] (RRID:SCR_018385) . This required downloading each pre-print as a PDF and then converting the PDFs to text files. 
<!-- Some papers from SocArXiv were only available in Word document format (.docx) and were downloaded as such and then converted to text format.  -->
We then conducted the open data and open code detection procedure, which involved searching for keywords and other markers of open data and open code availability. This was conducted using the `open_data_search()` function from the `oddpub` package. Under original validation process, the ODDPub algorithm performed with a sensitivity of 0.73 and a specificity of 1.00 for open code detection and a sensitivity of 0.73 and a specificity of 0.97 for open data detection compared with manual screening [@citePDDPub].

What constitutes open code or open data is complicated and discipline specific. The details of the `oddpub` approach are available in @citePDDPub. The general criteria are that specific mention should be made of where the data and code are located, and that data should be as close to raw as possible. Data and code must also be freely accessible to anyone (no request, application, registration process, or affiliation required). Our work was conducted using the statistical programming language R [@citeR] (RRID:SCR_001905).

@@

<!-- Although SocArXiv's metadata contained author-reported information concerning data and code availability for each pre-print, we elected to use the results of the ODDPub algorithm in our analysis to ensure comparability of results between pre-prints from all repositories.  -->
<!-- Notable discrepancies between the ODDPub algorithm and authors' self-reported data and code availability are documented in Appendix \@ref(socarxivissues). -->

```{r monthly-papers-total, fig.cap = "Number of pre-prints related to COVID-19 posted to arXiv (N = 4,250), bioRxiv (N = 4,146), and medRxiv (N = 13,194) per month", out.width="90%", fig.align = "center", fig.pos="H"}
just_arxiv <- 
  all_arxiv_results  %>% 
  select(submitted) %>% 
  mutate(type = "arXiv") %>% 
  rename(date = submitted)

@@

# just_socarxiv <- 
#   socarxiv_open_data_results  %>% 
#   select(date_created) %>%
#   rename(date = "date_created") %>%
#   mutate(date = as.Date(date), type = "SocArXiv")

all <- rbind(just_arxiv, 
             just_medRxiv, 
             just_bioRxiv
             # , 
             # just_socarxiv
             )

@@

```{r 2019-results-med-arxiv, out.width="90%", fig.align='center', fig.cap = "Number of pre-prints posted per month in 2019 to arXiv (N = 150,018), bioRxiv (N = 31,752), medRxiv (N = 913)", fig.pos="H"}
bio_2019 <- bio_2019_results %>%
  select(date) %>%
  mutate(type = "bioRxiv")

@@

# socarxiv_2019 <- socarxiv_metadata_2019 %>%
#   select(date_created) %>%
#   mutate(type = "socArXiv") %>%
#   rename(date = date_created)

@@

all_2019 <- rbind(arxiv_2019, bio_2019, med_2019
                  # , socarxiv_2019
                  )

@@

Between June and December 2019, the number of pre-prints posted to medRxiv monthly saw an overall increase, which may be expected as the repository gained recognition and popularity in the medical research community (Figure \@ref(fig:2019-results-med-arxiv)). The number of pre-prints posted monthly to bioRxiv also saw a slight overall increase throughout 2019, while the number of those posted to arXiv fluctuated throughout the year (Figure \@ref(fig:2019-results-med-arxiv)). Due to its relative immaturity at the beginning of the COVID-19 pandemic, a significant portion of medRxiv's overall usage has been dedicated to COVID-19-related research. In total, 21,647 pre-prints were posted to medRxiv between June 2019 and 30 June, 2021, 13,194 of which (approximately 61 per cent) relate to COVID-19.

Of the analyzed pre-prints from 2019, 93 per cent of those posted to arXiv, 63 per cent of those posted to bioRxiv, and 75 per cent of those posted to medRxiv showed no indication of open data or open code (Table \@ref(tab:2019-published)).

```{r monthly-papers-condition-stack-2019, fig.cap = "Number of sampled pre-prints posted in 2019 from arXiv (n = 1,200), bioRxiv (n = 1,200), and medRxiv (n = 913) distinguished by presence of open data or code markers", out.width="90%", fig.align = "center"}
# Number of papers for each open data/code status per month
open_status_med_2019 <-
  med_2019_open_data_results %>%
  select(title, date, is_open_code, is_open_data) %>%
  mutate(condition = is_open_data + is_open_code, repo = "medRxiv")

@@

# open_status_socarxiv_2019 <-
#   socarxiv_2019_open_data_results %>%
#   select(title, date_created, is_open_code, is_open_data) %>%
#   mutate(condition = is_open_data + is_open_code, repo = "SocArXiv", date = as.character(date_created)) %>%
#   select(title, date, is_open_code, is_open_data, condition, repo)

open_status <- rbind(open_status_med_2019, open_status_bio_2019, open_status_arxiv_2019
                     # , open_status_socarxiv_2019
                     )

@@

figufour <- 
open_status %>%
  ggplot() +
  geom_bar(aes(x = factor(floor_date(date, "month")),
              fill=factor(condition, levels = c("Neither", "Open Code", "Open Data", "Both")),
              position = "stack")) +
  labs(fill = "Type of marker") +
  scale_x_discrete(breaks=c("2019-01-01","2019-07-01", "2019-12-01"),
                   labels=c("Jan 2019", "Jul 2019", "Dec 2019")) +
  ylab("Number of pre-prints") +
  xlab("Month") +
  labs(title="Sampled pre-prints by month in 2019", subtitle="Note the different y-axes.") +
  theme_minimal() +
  scale_fill_manual(values = cbPalette) +
  facet_wrap(~repo,
             scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

@@

# socarxiv_published_2019 <- 
#   socarxiv_metadata_2019 %>% 
#   select(title, date_created, links.doi) %>% 
#   mutate(repo = "socArXiv", published = as.numeric(!is.na(links.doi))) %>%
#   rename(date = date_created) %>%
#   select(title, date, published, repo)

@@

published_2019 <- rbind(med_published_2019, bio_published_2019, 
                        # socarxiv_published_2019, 
                        arxiv_published_2019)

@@

# socarxiv_2019_published_open_status <- 
#   socarxiv_2019_open_data_results %>%
#   mutate(published = !is.na(links.doi) %>% as.numeric()) %>%
#   select(title, date_created, published, is_open_code, is_open_data) %>%
#   rename(date = date_created) %>%
#   mutate(condition = is_open_data + is_open_code, repo = "socArXiv")

@@

published_2019_open_status <- rbind(med_2019_published_open_status, bio_2019_published_open_status, 
                                    # socarxiv_2019_published_open_status, 
                                    arxiv_2019_published_open_status)

@@

published_2019_open_status %>%
  count(repo, published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>%
  mutate(Status = ifelse(published == 0, "Unpublished", "Published"), `Proportion with neither`= round((`Neither`)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  select(Status, Both, Neither, `Open code`, `Open data`, `Proportion with neither`) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published, 2019",
        booktabs = TRUE,
        format.args = list(big.mark = ",")) %>%
  group_rows(index = c("arXiv" = 2, "bioRxiv" = 2, "medRxiv" = 2))
```

Examining publication rates for pre-pandemic papers, we observe that 41 per cent of pre-prints posted to arXiv, 64 per cent of pre-prints posted to bioRxiv, and 61 per cent of pre-prints posted to medRxiv during 2019 were eventually peer reviewed and published (Table \@ref(tab:2019-published)). When disaggregated by open data and code status, we find that published and unpublished pre-prints contain open data and code markers in similar proportions (Table \@ref(tab:2019-open-publishing)).

@@

From the collection of all pre-prints related to COVID-19, we randomly sampled 3,910 pre-prints to analyze, stratified by repository. This sample is broken down as follows: 1,500 from medRxiv, 1,000 from arXiv, and 1,000 from bioRxiv. Broadly, we are unable to find markers of either open data or open code for 3,011 pre-prints or approximately 76 per cent of our sample (Table \@ref(tab:summarycounts-repositories)). Of the remaining pre-prints, 7 per cent contained open code markers only, 10 per cent contained open data markers only, and 8 per cent included markers of both open data and open code.

When differentiated by repository, we observe that open data and code markers were absent from 75 per cent of the sampled arXiv pre-prints, 67 per cent of the sampled bioRxiv pre-prints, 79 per cent of the sampled medRxiv pre-prints. The distribution of the remaining portion of pre-prints also varies by repository (Table \@ref(tab:summarycounts-repositories)). Notably, 28 per cent of sampled pre-prints from bioRxiv contained open data markers and 22 per cent of sampled arXiv pre-prints contained markers of open code, the highest proportions of any repository for each type of marker. Our results are similar to @mcguinness2021descriptive, who focus on medRxiv and find that 23 per cent describe open data.

@@

# just_socarxiv <-
#   socarxiv_open_data_results  %>%
#   select(date_created, is_open_data, is_open_code) %>%
#   rename(date = "date_created") %>%
#   mutate(type = "SocArXiv")

all <- rbind(just_arxiv, just_medRxiv, just_bioRxiv
             # , just_socarxiv
             )

@@

The distribution of total sampled pre-prints and sampled pre-prints with open data or code markers roughly follows that of COVID-19-related pre-prints posted in general (Figure \@ref(fig:arxiv-and-monthly-papers-condition-stack)). The proportion of pre-prints with open data or code has fluctuated over time but shows no consistent overall increase or decrease throughout the course of the pandemic, nor in conjunction with increases or decreases in the total number of pre-prints posted to any given repository. In our datasets, very few (if any) pre-prints were sampled for the month of January 2020. None of these pre-prints contained open data or open code markers, thus the 0 per cent rate of open data and code for this month across all repositories should be considered an outlier.

It is also important to note that pre-prints posted during the early months of the pandemic were likely using, and reusing, publicly available data sources due to an inability to collect original data within a short timeframe. Additionally, `oddpub` does not consider '[t]he reuse of data/code previously published by other researchers' [@citePDDPub]. A different definition of open data could enable pre-prints that reuse publicly available data to have made theirdata available for reproducibility purposes.

@@

# Tables with specific values displayed in the graphs

@@

all_2019 <- rbind(just_2019_medRxiv, just_2019_bioRxiv, 
                  # just_2019_socarxiv, 
                  just_2019_arxiv)
